They're Clutching up! Team Momentum in Round-Based Esports

### 2022-08-15 paper/paper.Rmd
rm little p
more prose

```{r}
#| label: cod_round_win_prop_after_x_wins
## http://keyonvafa.com/hot-hand/
get_post_streak_prob <- function(n, k, p = 0.5) {
  tosses <- rbinom(n, 1, p)
  runs <- rle(tosses)
  n_neg_after <- length(which(runs$values == 1 & runs$lengths >= k))
  n_pos_after <- sum(runs$lengths[which(runs$values == 1 & runs$lengths >= k)] - k)

@@

summarize_cod_win_prop_after_x_wins <- function(k) {
  probs_nx_k <- crossing(
    n_rounds = 7L:11L,
    win_series = c(TRUE, FALSE)
  ) |> 
    mutate(
      is_valid = case_when(
        win_series ~ TRUE,
        (n_rounds - !!k - 6L) > 0L ~ TRUE,
        TRUE ~ FALSE
      )
    ) |> 
    filter(is_valid) |> 
    select(-is_valid) |> 
    mutate(
      prior_p = (ifelse(win_series, 6L, n_rounds - 6L) - k) / (n_rounds - k),
      expected_poststreak_p = map2_dbl(
        n_rounds, prior_p, 
        ~simulate_post_streak_prob(n = ..1, k = !!k, p = ..2)
      )
    )

@@

cod_round_win_prop_after_x_wins <- tibble(win_streak = 2L:4L) |>
  mutate(
    data = map(win_streak, summarize_cod_win_prop_after_x_wins)
  ) |>
  unnest(data)
```

```{r}
#| label: cod_round_win_prop_after_x_wins-table
cod_round_win_prop_after_x_wins |> 
  mutate(
    across(
      matches('p_value'), 
      ~case_when(
        .x <= 0.01 ~ '<=0.01', 
        round(.x, 2) == 1 ~ '1',
        TRUE ~ as.character(round(.x, 2))
      )
    )
  ) |>
  # filter(win_streak == 3) |> 
  select(
    win_streak,
    win_series,
    n_rounds,
    total,
    prop,
    prior_p,
    expected_poststreak_p,
    gt_p_value,
    gt_p_value_adj
  ) |> 
  gt::gt() |> 
  gt::fmt_percent(
    decimals = 0,
    columns = c(prop, prior_p, expected_poststreak_p)
  ) |> 
  gt::cols_label(
    win_streak = 'Streak',
    win_series = 'Win Series?',
    n_rounds = 'Series lasts $i$ rounds',
    total = 'N',
    prop = 'Obs.',
    prior_p = 'Naive',
    expected_poststreak_p = 'Discounted',
    gt_p_value = 'Naive',
    gt_p_value_adj = 'Discounted'
  ) |> 
  gt::tab_spanner(
    label = '%s',
    columns = c(prop, prior_p, expected_poststreak_p)
  ) |> 
  gt::tab_spanner(
    label = 'Binom. > p-value',
    columns = c(gt_p_value, gt_p_value_adj)
  )
```

@@

```{=tex}
\begin{equation}\protect\hypertarget{eq:pki}{}{
  \hat{P}(\text{win round | k-round winning streak, series lasting i rounds})=\{
  \begin{array}{cl}
    \frac{6 - k}{i - k}, & \text{team wins series} \\
    \frac{11 - 6 - k}{i - k}, & \text{team loses series}
  \end{array}.
}\label{eq:pki}\end{equation}
```
Miller et al. [-@miller2018] provide evidence for negative recency suggesting that

```{=tex}
\begin{equation}\protect\hypertarget{eq:pkid}{}{
  \begin{array}{rcl}
  \hat{P}_{ki,\text{diff}} & = & \hat{P}(\text{win round | k-round winning streak, series lasting i rounds}) -  \\
    & & \hat{P}(\text{lose round | k-round losing streak, series lasting i rounds}) \\
    E[\hat{P}_{ki,\text{diff}}] & \neq & 0
  \end{array}
}\label{eq:pkid}\end{equation}
```
given a deterministic i.i.d. Bernoulli probability $\hat{p}(\text{win round | k-round winning streak, series lasting i rounds)}$ (Equation \ref{eq:pki}.[^13]

[^13]: Note that these probabilites are separate from those in Section \\ref{sec:analysis-1}.

Given that people typically perceive streaks as beginning after the third success (or failure) at minimum [@carlson2007], I focus on round win percentages in rounds immediately following streaks of three wins, $k = 3$.[^14]

@@

```{r}
#| label: cod_round_win_prop_after_x_wins_and_rounds
summarize_cod_win_prop_after_x_wins_y_rounds <- function(k) {
  probs_nx_k <- crossing(
    n_rounds = 7L:11L,
    round = 3L:11L
  ) |> 
    filter(round >= (k + 1L)) |> 
    mutate(
      prior_p = 6L / n_rounds,
      expected_poststreak_p = map2_dbl(
        n_rounds, prior_p, 
        ~simulate_post_streak_prob(n = ..1, k = !!k, p = ..2)
      )
    )

  postprocess_round_streaks(
    k = k,
    probs_nx_k = probs_nx_k,
    round
  )
}

cod_round_win_prop_after_x_wins_y_rounds <- tibble(win_streak = 2L:4L) |>
  mutate(
    data = map(win_streak, summarize_cod_win_prop_after_x_wins_y_rounds)
  ) |>
  unnest(data)
```

>>>>

```{r}
#| label: cod_round_win_prop_after_k_wins
## http://keyonvafa.com/hot-hand/
get_post_streak_prob <- function(n, k, p = 0.5) {
  tosses <- rbinom(n, 1, p)
  runs <- rle(tosses)
  n_neg_after <- length(which(runs$values == 1 & runs$lengths >= k))
  n_pos_after <- sum(runs$lengths[which(runs$values == 1 & runs$lengths >= k)] - k)

@@

summarize_cod_win_prop_after_k_wins <- function(k) {
  probs_nx_k <- crossing(
    n_rounds = 7L:11L,
    win_series = c(TRUE, FALSE)
  ) |> 
    mutate(
      is_valid = case_when(
        win_series ~ TRUE,
        (n_rounds - !!k - 6L) > 0L ~ TRUE,
        TRUE ~ FALSE
      )
    ) |> 
    filter(is_valid) |> 
    select(-is_valid) |> 
    mutate(
      prior_p = (ifelse(win_series, 6L, n_rounds - 6L) - k) / (n_rounds - k),
      expected_poststreak_p = map2_dbl(
        n_rounds, prior_p, 
        ~simulate_post_streak_prob(n = ..1, k = !!k, p = ..2)
      )
    )

@@

cod_round_win_prop_after_k_wins <- tibble(win_streak = 2L:4L) |>
  mutate(
    data = map(win_streak, summarize_cod_win_prop_after_k_wins)
  ) |>
  unnest(data)
```

```{r}
#| label: cod_round_win_prop_after_k_wins-table
cod_round_win_prop_after_k_wins |> 
  mutate(
    across(
      matches('p_value'), 
      ~case_when(
        .x <= 0.01 ~ '<=0.01', 
        round(.x, 2) == 1 ~ '1',
        TRUE ~ as.character(round(.x, 2))
      )
    )
  ) |>
  # filter(win_streak == 3) |> 
  select(
    win_streak,
    win_series,
    n_rounds,
    total,
    prop,
    prior_p,
    expected_poststreak_p,
    gt_p_value,
    gt_p_value_adj
  ) |> 
  gt::gt() |> 
  gt::fmt_percent(
    decimals = 0,
    columns = c(prop, prior_p, expected_poststreak_p)
  ) |> 
  gt::cols_label(
    win_streak = 'Streak',
    win_series = 'Win Series?',
    n_rounds = 'Series lasts $i$ rounds',
    total = 'N',
    prop = 'Obs.',
    prior_p = 'Naive',
    expected_poststreak_p = 'Discounted',
    gt_p_value = 'Naive',
    gt_p_value_adj = 'Discounted'
  ) |> 
  gt::tab_spanner(
    label = '%s',
    columns = c(prop, prior_p, expected_poststreak_p)
  ) |> 
  gt::tab_spanner(
    label = 'Binom. > p-value',
    columns = c(gt_p_value, gt_p_value_adj)
  )
```

@@

```{=tex}
\begin{equation}\protect\hypertarget{eq:pki}{}{
  \hat{P}_{wki} = \hat{P}(\text{win round | k-round winning streak, series lasting i rounds})=\{
  \begin{array}{cl}
    \frac{6 - k}{i - k}, & \text{team wins series} \\
    \frac{11 - 6 - k}{i - k}, & \text{team loses series}
  \end{array}.
}\label{eq:pki}\end{equation}
```
$\hat{P}_{lki} = \hat{P}(\text{lose round | k-round losing streak, series lasting i rounds})$ can be formulated symmetrically.

Miller et al. [-@miller2018] provide evidence for negative recency suggesting that

```{=tex}
\begin{equation}\protect\hypertarget{eq:pkid}{}{
  \begin{array}{rcl}
  \hat{P}_{ki,\text{diff}} & = & \hat{P}(\text{win round | k-round winning streak, series lasting i rounds}) -  \\
    & & \hat{P}(\text{lose round | k-round losing streak, series lasting i rounds}) \\
    E[\hat{P}_{ki,\text{diff}}] & \neq & 0
  \end{array}
}\label{eq:pkid}\end{equation}
```
given a deterministic i.i.d. Bernoulli probability (Equation \ref{eq:pki}.[^13]

[^13]: Note that these probabilities are separate from those in Section \ref{sec:analysis-1}.

Given that people typically perceive streaks as beginning after the third success (or failure) at minimum [@carlson2007], I focus on round win percentages in rounds immediately following streaks of three wins.[^14] Table \ref{tbl:cod-round-win-prop-after-k-wins} shows $\hat{P}_{wki}$ for $k = 3$ winning streaks, for both teams that empirically won and lost the series. (Some combinations of $i$ and $\text{team loses series}$ are irrelevant, such as the case when $i \in [7, 8, 9]$, so these combinations are excluded.)

@@
